---
title: "Introduction to rjags: A Bayesian linear regression model"
output: html_document
date: "Jan-30-2025"
---
The root directory for this project has 3 folders: "data", "src" and "run". The source Rmd file (this file) is in the "run" directory. The "data" directory contains a dataset from the National Cancer Institute's Surveillance, Epidemiology and End Results (SEER) database. This report presents a Bayesian linear regression analysis of the Stockton real estate dataset. The objective is to model the relationship between the outcome variable **Y** (home values) and covariates **X1** through **X6** using Bayesian methods implemented via JAGS. The dataset is analyzed to identify significant predictors of home values and assess the model's predictive performance.
 The "src" directory includes a file "modelJags.txt" that encodes a Bayesian linear regression model in the JAGS language and an R program called "dataPrep.r" to prepare the data for the JAGS model. 

1. Remove all objects from the R workspace---good programming practice and can help debug your code better.
```{r}
rm(list = ls())
```

2. Load "rjags" and "coda" packages
```{r}
library(rjags)
library(coda)
```

3. Prepare the data by running the "dataPrep.r" file in the "src" directory. 
```{r}
source("../src/dataPrep.r")
```

4. We initialize the unknown parameters in the JAGS model. 
```{r}
inits = list(
                list(beta = rep(0, times=p), tausq=1.0, yFit = rep(0, times=N))
)
```

5. Set the following parameters for Markov chain Monte Carlo (MCMC):
```{r}
nSamp = 10000 ## Number of posterior samples for inference
nChains = 1 ## Number of chains
nAdapt = 200 ## The initial number of runs to tune parameters
nBurn = 1000 ## Number of initial runs ("burn" period) for MCMC to converge
```

6. Define a set of model parameters for JAGS to recognize
```{r}
modelParameters = c("beta", "tausq", "sigmasq", "yFit")
```

7. Compile the Bayesian model
```{r}
m <- jags.model("../src/modelJags.txt", data=jagsData, inits=inits, n.chains=nChains, n.adapt=nAdapt)
```

8. Update the model for the burn-in period
```{r, message=FALSE, warning=FALSE}
update(m, n.iter=nBurn)
```

9. Sample from the posterior distribution for inference
```{r, message=FALSE, warning=FALSE}
mOut <- coda.samples(model=m, variable.names=modelParameters, n.iter=nSamp)
```

10. Get descriptive statistics for posterior distributions(posterior mean, posterior standard deviation and 95% credible intervals for: (i) the regression slopes; (ii) the residual variance; and (iii) the model-fitted values of the data)
```{r}
sm <- summary(mOut)
sm
```

11. Extract samples of all monitored parameters from the full posterior distribution
```{r}
samps <- do.call(rbind, mOut)
```

12. Medians and 95% credible intervals
```{r}
credibleIntervals <- t(apply(samps, 2, function(x){quantile(x, c(0.50, 0.025, 0.975))}))
ciBeta <- credibleIntervals[grep("beta", rownames(credibleIntervals)),]
ciBeta
ciyFit <- credibleIntervals[grep("yFit", rownames(credibleIntervals)),]
ciyFit
```


```{r}
posterior_median <- apply(samps, 2, median)
posterior_median

```

14. Which of the covariates are statistically significant predictors of Y
```{r}
significant_covariates <- rownames(ciBeta)[!(ciBeta[,2] < 0 & ciBeta[,3] > 0)]
significant_covariates

```

15.How many of the 95% credible intervals for the model-fitted values actually include the observed values?
```{r}
included_observed <- sum(data$Y >= ciyFit[,2] & data$Y <= ciyFit[,3])
included_observed

```

16. Scatter plot of observed vs. fitted median values
```{r}
# Extracting the median fitted values
head(ciyFit) 
yFitMedians <- ciyFit[, 1]  # First column contains the medians

# Scatter Plot: Observed vs. Fitted Median Values
plot(Y, yFitMedians, 
     main = "Observed vs. Fitted Median Values", 
     xlab = "Observed Y", 
     ylab = "Fitted Median Y", 
     pch = 19, 
     col = "blue")

# Add a 45-degree reference line
abline(0, 1, col = "red", lwd = 2, lty = 2)

# Calculate and Display Correlation
correlation_result <- cor(Y, yFitMedians)

# Adding correlation value to the plot
legend("topleft", legend = paste("Correlation =", round(correlation_result, 3)), 
       bty = "n", text.col = "darkred", cex = 0.9)

# Display correlation value in console as well
print(paste("Correlation between Observed Y and Fitted Median Y:", round(correlation_result, 3)))

```