---
title: "Introduction to rjags: A Bayesian linear regression model"
output: html_document
date: "Feb-28-2025"
---
0. The root directory for this project has 3 folders: "data", "src" and "run". The source Rmd file (this file) is in the "run" directory. The "data" directory contains the real estate dataset from Stockton, which includes housing attributes. We will use a Bayesian linear regression model to predict housing prices based on these attributes. The "src" directory includes a file "modelJagsPredictive.txt" that encodes a Bayesian linear regression model in the JAGS language and an R program called "dataPrepPredictive.r" to prepare the data for the JAGS model. 

1. Remove all objects from the R workspace---good programming practice and can help debug your code better.
```{r}
rm(list = ls())
```

2. Load "rjags" and "coda" packages
```{r}
library(rjags)
library(coda)
set.seed(1234)
```

3. Prepare the data by running the "dataPrep.r" file in the "src" directory. 
```{r}
source("../src/dataPrepPredictive.r")

```

4. We initialize the unknown parameters in the JAGS model. 
```{r}
inits = list(
                list(beta = rep(0, times=p), tausq=1.0, yFit = rep(0, times=N))
)
```

5. Set the following parameters for Markov chain Monte Carlo (MCMC):
```{r}
nSamp = 10000 ## Number of posterior samples for inference
nChains = 1 ## Number of chains
nAdapt = 200 ## The initial number of runs to tune parameters
nBurn = 1000 ## Number of initial runs ("burn" period) for MCMC to converge
```

6. Define a set of model parameters for JAGS to recognize
```{r}
modelParameters = c("beta", "tausq", "sigmasq", "yFit", "yPred")

```

7. Compile the Bayesian model
```{r}
m <- jags.model("../src/modelJagsPredictive.txt", data=jagsData, inits=inits, n.chains=nChains, n.adapt=nAdapt)
```

8. Update the model for the burn-in period
```{r, message=FALSE, warning=FALSE}
update(m, n.iter=nBurn)
```

9. Sample from the posterior distribution for inference
```{r, message=FALSE, warning=FALSE}
mOut <- coda.samples(model=m, variable.names=modelParameters, n.iter=nSamp)
```

10. Get descriptive statistics for posterior distributions
```{r}
sm <- summary(mOut)
```

11. Extract samples of all monitored parameters from the full posterior distribution
```{r}
samps <- do.call(rbind, mOut)
```

12. Medians and 95% credible intervals
```{r}
credibleIntervals <- t(apply(samps, 2, function(x){quantile(x, c(0.50, 0.025, 0.975))}))
ciBeta <- credibleIntervals[grep("beta", rownames(credibleIntervals)),]
ciBeta
ciyFit <- credibleIntervals[grep("yFit", rownames(credibleIntervals)),]
ciyFit
```
